PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
fixed Highest education prepandemic

## Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# NBR demographics
## Read in data
```{r nbr dem cols}
if(NBR == TRUE){
  nbr.dem.raw <- readRDS(file = paste0(data.raw_path, "/nbr/dem_nbr.rds"))
  dim(nbr.dem.raw)
  colnames(nbr.dem.raw)
}
```

## Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "Age_uncleaned",
                        "EduYrs")

@@

## Cohort information
Information on INTERVAL, COMPARE, and STRIDES
http://www.donorhealth-btru.nihr.ac.uk/studies/
```{r nbr cohort information}
if(NBR == TRUE){
  nbr.cohort.raw <- readxl::read_xlsx(
    path = paste0(data.raw_path, "/nbr/coping_participants_export_20200730.xlsx")
    )

@@

```{r}

@@

#Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# define age limits 
age_upper_limit = 117 # oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect colnames
colnames(dem.raw.id)
}
```

@@

#recode the numeric version into a factor
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

>>>>

### Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## NBR
### Read in data
```{r nbr dem cols}
if(NBR == TRUE){
  nbr.dem.raw <- readRDS(file = paste0(data.raw_path, "/nbr/dem_nbr.rds"))
  dim(nbr.dem.raw)
  colnames(nbr.dem.raw)
}
```

### Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "Age_uncleaned",
                        "EduYrs")

@@

### Cohort information
Information on INTERVAL, COMPARE, and STRIDES
http://www.donorhealth-btru.nihr.ac.uk/studies/
```{r nbr cohort information}
if(NBR == TRUE){
  nbr.cohort.raw <- readxl::read_xlsx(
    path = paste0(data.raw_path, "/nbr/coping_participants_export_20200730.xlsx")
    )

@@

@@

#Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16

@@

# Inspect variable
dem.raw.id %>%
  descr(
    Age,
    #stats = "common"
    )
}
```

@@

#recode the numeric version into a factor
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan =
      recode_factor(
        highest_education_prepan_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )
